Inference methods comparison for the simplified Naomi model using the Kolmogorov-Smirnov test statistic

### 2023-05-24 src/docs_paper/paper.Rmd
Example KS plot into manuscript

The two-sample Kolmogorov-Smirnov (KS) test statistic [@smirnov1948table] is the maximum absolute difference between two empirical cumulative distribution (ECDF) functions $F(\varphi) = \frac{1}{n} \sum_{i = 1}^n \mathbb{I}_{\phi_i \leq \varphi}$.
We compare the KS statistics
\begin{align*}
D_\text{TMB} &= \sup_\varphi | F_\text{NUTS}(\varphi) - F_\text{TMB}(\varphi)|, \\
D_\text{PCA-AGHQ} &= \sup_\varphi | F_\text{NUTS}(\varphi) - F_\text{PCA-AGHQ}(\varphi)|.
\end{align*}
See a summary of the results in Table.

```{r ks, fig.cap="Results of Kolmogorov-Smironv tests analysis."}
ks_summary <- readRDS("depends/ks_summary.rds")

>>>>

The two-sample Kolmogorov-Smirnov (KS) test statistic [@smirnov1948table] is the maximum absolute difference between two empirical cumulative distribution (ECDF) functions $F(\varphi) = \frac{1}{n} \sum_{i = 1}^n \mathbb{I}_{\phi_i \leq \varphi}$.
We compare the KS statistics
\begin{align*}
D_\text{TMB} &= \sup_\varphi | F_\text{NUTS}(\varphi) - F_\text{TMB}(\varphi)|, \\
D_\text{PCA-AGHQ} &= \sup_\varphi | F_\text{NUTS}(\varphi) - F_\text{PCA-AGHQ}(\varphi)|.
\end{align*}
See an illustration of the KS test in Figure \ref{fig:ks} and a summary of the results in Table.

```{r ks, fig.height = 3, fig.cap="Example KS test for one parameter."}
df_compare <- readRDS("depends/beta_alpha.rds") %>%
  filter(method != "adam")

colours <- c("#56B4E9", "#009E73", "#E69F00")

mean <- df_compare %>%
  filter(method == "tmbstan") %>%
  summarise(mean = mean(samples)) %>%
  pull(mean) %>%
  round(digits = 3)

sd <- df_compare %>%
  filter(method == "tmbstan") %>%
  summarise(sd = sd(samples)) %>%
  pull(sd) %>%
  round(digits = 3)

histogram <- ggplot(df_compare, aes(x = samples, fill = method, col = method)) +
  geom_histogram(aes(y = after_stat(density)), alpha = 0.5, position = "identity", bins = 30) +
  theme_minimal() +
  facet_grid(method~.) +
  labs(x = "beta_alpha", y = "Density", fill = "Method") +
  scale_color_manual(values = colours) +
  scale_fill_manual(values = colours) +
  theme(legend.position = "none")

grid <- seq(from = min(df_compare$samples), to = max(df_compare$samples), length.out = 1000)

tmb_ecdf <- stats::ecdf(filter(df_compare, method == "TMB") %>% pull(samples))
tmb_ecdf_df <- data.frame(x = grid, ecdf = tmb_ecdf(grid), method = "TMB")

aghq_ecdf <- stats::ecdf(filter(df_compare, method == "aghq") %>% pull(samples))
aghq_ecdf_df <- data.frame(x = grid, ecdf = aghq_ecdf(grid), method = "aghq")

tmbstan_ecdf <- stats::ecdf(filter(df_compare, method == "tmbstan") %>% pull(samples))
tmbstan_ecdf_df <- data.frame(x = grid, ecdf = tmbstan_ecdf(grid), method = "tmbstan")

# Add ECDF differences
tmb_ecdf_df$ecdf_diff <- tmbstan_ecdf_df$ecdf - tmb_ecdf_df$ecdf
aghq_ecdf_df$ecdf_diff <- tmbstan_ecdf_df$ecdf - aghq_ecdf_df$ecdf
tmbstan_ecdf_df$ecdf_diff <- 0

absmax <- function(x) x[which.max(abs(x))]

ks_tmb <- absmax(tmb_ecdf_df$ecdf_diff)
ks_aghq <- absmax(aghq_ecdf_df$ecdf_diff)

ecdf_df <- bind_rows(tmb_ecdf_df, aghq_ecdf_df, tmbstan_ecdf_df)

ecdf_df$method <- factor(ecdf_df$method, levels = c("TMB", "aghq", "tmbstan"))

ks_labeller <- function(x) toString(round(abs(x), 2))

ecdf_diff <- ggplot(ecdf_df, aes(x = x, y = ecdf_diff, col = method)) +
  geom_line() +
  geom_abline(intercept = ks_tmb, slope = 0, col = colours[1], linetype = "dashed", alpha = 0.8) +
  annotate("text", x = 1.1 * max(ecdf_df$x), y = ks_tmb, label = ks_labeller(ks_tmb), col = colours[1], alpha = 0.8) +
  geom_abline(intercept = ks_aghq, slope = 0, col = colours[2], linetype = "dashed", alpha = 0.8) +
  annotate("text", x = 1.1 * max(ecdf_df$x), y = ks_aghq, label = ks_labeller(ks_aghq), col = colours[2], alpha = 0.8) +
  scale_color_manual(values = colours) +
  labs(x = "beta_alpha", y = "ECDF difference") +
  guides(col = "none") +
  coord_cartesian(xlim = c(min(ecdf_df$x), max(ecdf_df$x)), clip = "off") +
  theme_minimal() +
  theme(plot.margin = unit(c(1, 3, 1, 1), "lines"))

histogram + ecdf_diff
```

```{r}
ks_summary <- readRDS("depends/ks_summary.rds")

### 2023-05-24 src/naomi-simple_ks/ks.Rmd
Example KS plot into manuscript

```{r}
pars_eval <- pars %in% c("beta_rho", "beta_alpha", "beta_anc_rho", "u_rho_x")
names(pars_eval) <- pars
```

@@

```{r eval = pars_eval[["beta_alpha"]]}
beta_alpha <- histogram_and_ecdf("beta_alpha", return_df = TRUE)
beta_alpha$plot

#' Save off KS results as example
saveRDS(
  beta_alpha$df_compare,
  file = "beta_alpha.rds"
)
```

@@

```{r eval = pars_eval[["beta_anc_rho"]]}
beta_anc_rho <- histogram_and_ecdf("beta_anc_rho", i = 1, return_df = TRUE)
beta_anc_rho$plot

#' Save off KS results as example
saveRDS(
  beta_anc_rho$df_compare,
  file = "beta_anc_rho.rds"
)
```

>>>>

```{r}
pars_eval <- pars %in% c("beta_rho", "u_rho_x")
names(pars_eval) <- pars
```

@@

```{r eval = pars_eval[["beta_alpha"]]}
histogram_and_ecdf_helper("beta_alpha")
```

@@

```{r eval = pars_eval[["beta_anc_rho"]]}
histogram_and_ecdf_helper("beta_anc_rho")
```
